---
title: "Lab 1"
author: "Kolton Knight"
date: "2025-01-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in data from CSV file, then view, finally attach as default object.

```{r dataframe}
EPI_data <- read.csv('C:/Users/kolto/Documents/Data Analytics/epi2024results06022024.csv')
View(EPI_data)
attach(EPI_data)
```

Check and Filter NA

```{r NA}
NAs <- is.na(EPI.new) 
EPI.new.noNAs <- EPI.new[!NAs]
```

Summary Stats 2 ways

``` {r summary EPI}
summary(EPI.new)
fivenum(EPI.new, na.rm=TRUE)
```

Stem and Leaf Plot
```{r s&L EPI}
stem(EPI.new)
```

Histogram with sequence specified (Min, Max, Step), Prob = true plots 
probability densities so total area = 1

```{r hist EPI}
hist(EPI.new, seq(20., 80., 1.0), prob=TRUE)
```


Histogram with Lines Drawn *Gotta run everything together
``` {r hist lines EPI}
hist(EPI.new, seq(20., 80., 1.0), prob=TRUE)
lines(density(EPI.new, na.rm=TRUE, bw=1.))
rug(EPI.new)
```

Histogram with smoother line
```{r hist smooth line EPI}
hist(EPI.new, seq(20., 80., 1.0), prob=TRUE)
lines(density(EPI.new, na.rm=TRUE, bw="SJ"))
rug(EPI.new)
```

Fitting two normal distributions in histogram w smooth line
```{r two norm EPI}
hist(EPI.new, seq(20., 80., 1.0), prob=TRUE)
lines(density(EPI.new, na.rm=TRUE, bw="SJ"))
rug(EPI.new)
x<-seq(20,80,1)
q<- dnorm(x,mean=42, sd=5,log=FALSE)
#lines(x,q)
lines(x,.4*q)
q<-dnorm(x,mean=65, sd=5,log=FALSE)
lines(x,.12*q)
```

Cumulative Density Function

```{r CDF EPI}
plot(ecdf(EPI.new), do.points=FALSE, verticals=TRUE)
```

Quantile-Quantile

```{r Q-Q EPI}
qqnorm(EPI.new)
qqline(EPI.new)
```

Q-Q plot for normal distribution
``` {r Q-Q norm EPI}
qqplot(rnorm(250), EPI.new, xlab = "Q-Q plot for norm dsn")
qqline(EPI.new)
```

Q-Q plot for t distribution
``` {r Q-Q t EPI}
qqplot(rt(250, df = 5), EPI.new, xlab = "Q-Q plot for t dsn")
qqline(EPI.new)
```

Summary Stats 2 ways - SPI.new

``` {r summary SPI}
summary(SPI.new)
fivenum(SPI.new, na.rm=TRUE)
```
Check and Filter NA

```{r NA SPI}
NAs <- is.na(SPI.new) 
SPI.new.noNAs <- SPI.new[!NAs]
```

Stem and Leaf Plot - SPI.new
```{r s&L SPI}
stem(SPI.new.noNAs)
```

Histogram with Lines Drawn - SPI.new
``` {r hist lines SPI}
hist(SPI.new.noNAs, seq(-5., 110., 1.0), prob=TRUE)
lines(density(SPI.new.noNAs, na.rm=TRUE, bw=1.))
rug(SPI.new.noNAs)
```

Histogram with Smooth Lines Drawn - SPI.new
``` {r hist lines smooth SPI}
hist(SPI.new.noNAs, seq(-5., 110., 1.0), prob=TRUE)
lines(density(SPI.new.noNAs, na.rm=TRUE, bw='SJ'))
rug(SPI.new.noNAs)
```

Fitting uniform distribution in histogram w smooth line
```{r uniform SPI}
hist(SPI.new.noNAs, seq(-5., 110., 1.0), prob=TRUE)
#lines(density(SPI.new.noNAs, na.rm=TRUE, bw="SJ"))
rug(SPI.new.noNAs)
x<-seq(-5., 110., 1.0)
q<- dunif(x,min = 0, max = 100, log=FALSE)
lines(x,q)
```

Comparing EPI.new and SPI.new
``` {r b&w plot SPI}
boxplot(EPI.new, SPI.new.noNAs)
```

Cumulative Density Function - SPI

```{r CDF SPI}
plot(ecdf(SPI.new.noNAs), do.points=FALSE, verticals=TRUE)
```

Quantile-Quantile - SPi

```{r Q-Q SPI}
qqnorm(SPI.new.noNAs)
qqline(SPI.new.noNAs)
```

Q-Q plot for normal distribution - SPi
``` {r Q-Q norm SPI}
qqplot(rnorm(250), SPI.new.noNAs, xlab = "Q-Q plot for norm dsn")
qqline(SPI.new.noNAs)
```

Q-Q plot for t distribution - SPi
``` {r Q-Q t SPI}
qqplot(rt(250, df = 5), SPI.new.noNAs, xlab = "Q-Q plot for t dsn")
qqline(SPI.new.noNAs)
```

Summary Stats 2 ways - TBN.new

``` {r summary TBN}
summary(TBN.new)
fivenum(TBN.new, na.rm=TRUE)
sd(TBN.new, na.rm=TRUE)
```
Check and Filter NA

```{r NA TBN}
NAs <- is.na(TBN.new) 
TBN.new.noNAs <- TBN.new[!NAs]
```

Comparing EPI.new and TBN.new
``` {r b&w plot TBN}
boxplot(EPI.new, TBN.new.noNAs)
```

Histogram with Lines Drawn - TBN.new
``` {r hist lines TBN}
hist(TBN.new.noNAs, seq(-5., 110., 1.0), prob=TRUE)
lines(density(TBN.new.noNAs, na.rm=TRUE, bw=1.))
rug(TBN.new.noNAs)
```

Fitting uniform distribution in histogram w smooth line - TBN
```{r uniform TBN}
hist(TBN.new.noNAs, seq(-5., 110., 1.0), prob=TRUE)
lines(density(TBN.new.noNAs, na.rm=TRUE, bw="SJ"))
rug(TBN.new.noNAs)
x<-seq(-5., 110., 1.0)
q<- dnorm(x, mean = 52, sd = 30.2679, log=FALSE)
lines(x,q)
```
Cumulative Density Function - TBN.new

```{r CDF TBN}
plot(ecdf(TBN.new), do.points=FALSE, verticals=TRUE)
```

Quantile-Quantile - TBN.new

```{r Q-Q TBN}
qqnorm(TBN.new)
qqline(TBN.new)
```

Q-Q plot for normal distribution - TBN.new
``` {r Q-Q norm TBN}
qqplot(rnorm(250), TBN.new, xlab = "Q-Q plot for norm dsn")
qqline(TBN.new)
```

Q-Q plot for t distribution - TBN.new
``` {r Q-Q t TBN}
qqplot(rt(250, df = 5), TBN.new, xlab = "Q-Q plot for t dsn")
qqline(TBN.new)
```
